Wait-free Queues with Polygarithmic Step Complexity

### 2022-05-13 article/psuedocode.tex
fix lemma claims

\usepackage{amsmath,amssymb,amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

@@

\Function{void}{Propagate()}{}
\If{\bf{not} \Call{Refresh()}{}}
\State \Call{Refresh()}{} \cmt{Lemma Double Refresh}
\EndIf
\If{\tt{this} \bf{is not} \tt{root}}
\State \tt{parent.}\Call{Propagate()}{}
\EndIf
\EndFunction{Propagate}

@@

\Function{boolean}{Refresh()}{}
\State \tt{s= size}
%\If{\tt{n.blocks[h]!=null?}} \tt{h+=1}\EndIf
\State \tt{<new, np\sub{left}, np\sub{right}>= \Call{CreateBlock}{h}} \cmt{\tt{np\sub{left}, np\sub{right}} are the values read from the children's \tt{num\sub{propagated}} feild.}
\If{\tt{new.num==0}} \Return{\tt{true}} \cmt{The block contains nothing.}
\ElsIf{\tt{blocks.tryAppend(new, s)}}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{okcas}
\State \tt{\Call{CAS}{dir.super[np\sub{dir}], null, h+1}} \cmt{Write would work too.}
\State \tt{\Call{CAS}{dir.num\sub{propagated}, np\sub{dir}, np\sub{dir}+1}}
\EndFor
\State \tt{\Call{CAS}{size, h, h+1}}
\State \Return{\tt{true}}
\Else
\State \tt{\Call{CAS}{size, h, h+1}} \cmt{Even if another process wins, help to increase the size. The winner might have fallen sleep before increasing \tt{size}.}
\State \Return{ \tt{false}}
\EndIf
\EndFunction{Refresh}

@@

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{blocks[b].num\sub{enq}$\geq$i}.}
\Function{<int, int>}{IndexDeq}{\sl{int} b, \sl{int} i} \cmt{Returns the rank of $i$th dequeue in the $b$th block of the node, among the dequeues in the root.}
\If{\tt{this} \bf{is} \tt{root}}
\State\Return \tt{<b, i>}
\Else
\State \tt{dir= (parent.left==n)? left: right} \cmt{check if a left or a right child}
\State \tt{superBlock= parent.\Call{BSearch}{sum\sub{deq-dir}, i, super[blocks[b].group]-p, super[blocks[b].group]+p}} 
\Statex\cmt{superblock's group has at most $p$ difference with the value stored in \tt{super[]}.}
\If{\tt{dir {\keywordfont is} right}}
%\State \tt{i+= parent.blocks[superBlock-1].sum\sub{deq-right}}
\State \tt{i+= blocks[superBlock].sum\sub{deq-left}} \cmt{consider the dequeues from the right child}
\EndIf
\State \Return\tt{this.parent.}\Call{IndexDeq}{superBlock, i}
\EndIf
\EndFunction{Index}

@@

\begin{definition}
If \nf{n.blocks[i]==b}  we call \nf{i} the \emph{index} of block \nf{b}. Block \nf{b} is before block \nf{b$^\prime$} in node \nf{n} if and only if \nf{b}'s index is smaller than \nf{b$^\prime$}'s. Block \nf{b} is propagated to node \nf{n} or set \nf{S} if \nf{b} is in \nf{n.blocks} or \nf{S} or is a subblock of a block in \nf{n.blocks} or \nf{S}.
\end{definition}

@@

\begin{definition}
  Block \nf{b} is is in the \emph{Established(n, t)} if \nf{b} is subblock of a \nf{b$^\prime$} such that \nf{n.size} at time $t$ is greater than index of \nf{b$^\prime$} .
\end{definition}

\begin{lemma}[sizeProgress] \label{lem::sizeProgress}
\nf{n.size} is non-decreasing over time.
\end{lemma}
%\begin{proof}
%  Simply because \nf{n.head} is only incremented.
%\end{proof}

\begin{lemma}[headPosition] \label{lem::headPosition} The value read in Line \ref{prevLine}(\nf{h=n.size}) might be 1 bit behind the first empty block in the node.
\end{lemma}
%\begin{proof}
% Because at the end of every \nf{Refresh()} with block size greater than 0 (Lines 53,56) \nf{n.head} is incremented. Maybe some process goes to sleep before incrementing the head, but after sleeping if \nf{h} does not increase then \nf{CAS} in Line 52 is going to be failed and nothing is going to be appended to \nf{n.blocks}.
%\end{proof}

@@

\begin{lemma}[establishedOrder]\label{lem::establishedOrder}
  If  time $t<$ time $t^\prime$, then $Established(n, t)\subseteq Established(n, t^\prime)$.
\end{lemma}
%\begin{proof}
%Because blocks are only appended(not modified) with CAS to \nf{n.blocks[n.head]} and \nf{n.head} is non-decreasing.
%\end{proof}

\begin{lemma}[createBlock] \label{lem::createBlock}
  If \nf{n.CreateBlock(h, x)} is invoked at time $t$ and $b$ is the block returned, then $Established(n.end_{left}, t) \cup Established(n.end_{right}, t)/ Established(n, t) \subseteq b$ .
\end{lemma}
%\begin{proof}
% We prove the claim for the left child. Blocks in \nf{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} are all the new established operations at time $t$ by definition of Subblock. Line 70 is after $t$ and since the head is only increasing (Lemma~\ref{lem::headProgress}) the lemma holds. See Figure~\ref{fig::createBlock}. The right child is the same.
%\end{proof}

\begin{lemma}[trueRefresh] \label{lem::trueRefresh}
  Suppose \nf{Refresh(n)}'s \nf{TryAppend(new, s)} returns \nf{true}. Let $t$ be the time \nf{n.Refresh()} is invoked, blocks in $Established(n.end_{left}, t)$ and $Established(n.end_{right}, t)$ are in $Established(n, t)$ after \nf{TryAppend(new, s)}.
\end{lemma}
%\begin{proof}
%By Lemma~\ref{lem::createBlock} \nf{new} contains \nf{n}'s childrens' established blocks before Line 43 which is  appended to \nf{n.blocks} by \nf{CAS} in Line 48.
%\end{proof}

\begin{lemma}[falseRefresh] \label{lem:falseRefresh}
    If instance $r$ of \nf{n.Refresh()} returns \nf{false}, then there is another successful instance $r^\prime$ of \nf{n.Refresh()} that has performed a successful \nf{TryAppend(new, s)} with the same value \nf{s} read.
\end{lemma}
%\begin{proof}
%  If there is no other concurrent successful \nf{Refresh(n)} then \nf{Refresh(n)} would succeed in Line 48. So there is another \nf{Refresh(n)}, that has to \nf{CAS}ed successfully its block in \nf{n.blocks[h]} after Line 43 of \nf{Refresh(n)}. Otherwise the other \nf{Refresh(n)} should have read \nf{h$^\prime$>h} instead of \nf{h} for \nf{n.head}(Line 52).
%\end{proof}

\begin{lemma}[Double Refresh] \label{doubleRefresh}
  Consider two consecutive failed instances $r_1$,$r_2$ of \nf{n.Refresh()} by some process. Let $t_1$ be the time is $r_1$ invoked and $t_2$ be the time $r_2$ is invoked. After $r_2$'s \nf{TryAppend} we have $Established(n.end_{left}, t) \cup Established(n.end_{right}, t) \subseteq Established(n, t)$.
\end{lemma}
%\begin{proof}
%  
%
%If Line 35 (first Refresh which we call $R_1$) returns \nf{true}, the claim is held by Lemma~\ref{lem::trueRefresh}. If not, then there is another successful instance of \nf{Refresh()} $R_1^\prime$ by Lemma~\ref{lem:falseRefresh}. $R_1^\prime$ may or may not have propagated some subblocks of \nf{new} to \nf{n}. It is obvious that the \nf{new} constructed by the second Refresh in Line 36 contains the blocks in \nf{new} by $R_1$ which $R_1^\prime$ did not contain, since \nf{n.head} is only increasing (Maybe Lemma~\ref{lem::oldnewOrder}). If $R_2$ succeeds by Lemma~\ref{lem::trueRefresh} the claim holds. If not, it is deduced that \nf{n.blocks[h]} was not \nf{null} before $R_2$'s \nf{CAS}. Furthermore, \nf{n.blocks[h]} was \nf{null} before reading \nf{h} by $R_1$. So there is a successful \nf{Refresh()} after the read of \nf{h} in $R_1$ and before the \nf{CAS} of $R_2$. This \nf{Refresh()} contains all the new established operations before Line 35 (Maybe Lemma~\ref{lem::oldnewOrder}) and by Lemma~\ref{lem::trueRefresh} our claim holds. See Figure~\ref{}.
%\end{proof}

%\begin{figure}[hbt]
%  \center\includegraphics[width=3in]{pics/doublyrefresh.drawio.png}
%  \caption{$R_2^\prime$'s \nf{CAS} is executed after \nf{h1=n.head}.}
%\end{figure}

 Block \nf{new} is created of new established subblocks of children of \nf{n}(Lemma \ref{lem::createBlock}, Line 46). If \nf{CAS} in Line 48 succeeds then by Lemma~\ref{lem::trueRefresh} new established blocks will be in \nf{n}.

\begin{lemma}[Double Refresh] \label{doublyRefresh}
All operations in \nf{n}'s children's blocks before line \nf{35} are guaranteed to be in \nf{n}'s blocks after Line~\nf{37}.
\end{lemma}
%\begin{proof}
%Suppose block \nf{b} with index \nf{i} is in the the left child of \nf{n} before the line 35. By Lemma~\ref{head} it follows that \nf{n.left.head} is greater than \nf{i}. \nf{Refresh()} calls \nf{CreateBlock()} and creates a block from blocks between \nf{n.blocks[n.head].end\textsubscript{left}} and \nf{n.left.head} in the left child, which contains \nf{b} as well. First it tries to append it in \nf{n.blocks.head} and if it was succsuful it continues recursivley. If not it tries again, and if the second call of \nf{Refresh()} in Line 36 fails. It means there is another \nf{Refresh} which has reah its \nf{i} after the Line 35, so it contains \nf{b} as well.
%\end{proof}
\nf{CreateBlock()} reads blocks in the children that do not exist in the parent and aggregates them into one block. If a \nf{Refresh()} procedure returns true it means it has appended the block created by \nf{CreateBlock()} into the parent node's sequence. So suppose two \nf{Refresh}es fail. Since the first \nf{Refresh()} was not successful, it means another CAS operation by a \nf{Refresh}, concurrent to the first \nf{Refresh()}, was successful before the second \nf{Refresh()}. So it means the second failed \nf{Refresh} is concurrent with a successful \nf{Refresh()} that assuredly has read block before the mentioned line \nf{35}. After all it means if any of the \nf{Refresh()} attempts were successful the claim is true, and also if both fail the mentioned claim still holds.

\begin{lemma}[Append]\label{append}
 When \text{\nf{Append(op)}} is finished, \nf{op} appears exactly once in a block of \nf{root.blocks}.
\end{lemma}
%\begin{proof}
%\nf{Append(op)} adds \nf{op} to \nf{l\textsubscript{p}.blocks}(Line~\ref{addOP}) and \nf{Propagate()} recursively propagates \nf{op} up to the root.
%By lemma \ref{doublyRefresh} we know that operation \nf{op} propagates from child to parent at each level.
%\end{proof}

\begin{lemma}[Block Size Upper Bound]\label{blockSize}
Each block in a node contains at most one operation from each processs.  
\end{lemma}
%\begin{proof}
%  Note that \nf{prevIndex} and \nf{lastIndex} defined in lines \ref{lastLine}, \ref{prevLine} are the indices defined in the Definition~\ref{def::subblock}. After a block has propagated to the parent the blocks between \nf{prevIndex} and \nf{lastIndex} make up to the parent (Lemma~\ref{doublyRefresh}). The number of new operations which have not propagated yet to the parent cannot be more than $p$. If so by the law of pigeonholes there is a process which has appended two concurrent operations.
%\end{proof}

\begin{lemma}[Subblocks Upperbound]\label{subBlocksBound}
Each block in a node  has at most $p$ subblocks.
\end{lemma}
%\begin{proof}
%  From Line~\ref{addOP} it is induced directly that each block contains at least one operation. Now it follows directly by Lemma \ref{blockSize}.
%\end{proof}

\begin{definition} [Ordering of operations inside a node] \label{ordering}
$\blacktriangleright$ Note that from Lemma \ref{blockSize} we know there is at most one operation from each process in a given block.

\begin{itemize}
  \item $E(n,i)$ is the sequence of enqueue operations that are member of \nf{n.blocks[i]} ordered by process id.
  \item 
$D(n,i)$ is the sequence of dequeue operations that are member of \nf{n.blocks[i]} ordered by process id.
\item $D(n)=D(n,1).D(n,2).D(n,3)...$
\item $L=E(root,1).D(root,1).E(root,2).D(root,2).E(root,3).D(root,3)...$
\end{itemize}
\end{definition}

@@

\begin{lemma}[Get] \label{get}
\nf{Get(n,b,i)} returns $i$th Enqueue in $E(n,b)$.
\end{lemma}
%\begin{proof}
%It is obvious that \nf{Get(leaf l,b,1)} returns the operations stored in $b$th block of leaf $l$. To find the \nf{i}th enqueue in block \nf{b} of an internal node \nf{n} in line 87 it is decided that it resides in the left child or the right child. This decision is made by Definition of $E(n,b)$. After that Lines 88, 92 search the proper subblocks of \nf{b}. From Definition~\ref{def::subblock} weknow the subblocks of the $b$th block are within the \nf{prevBlock} and \nf{lastBlock} block of the \nf{CreateBlock()}.
%\end{proof}

\begin{lemma}[Index]
 Index(n,b,i) returns the rank in the $D(root)$ of ith Dequeue in $D(n,b)$.
\end{lemma}
%\begin{proof}
%  \nf{Index(n,b,i)} computes superblock of $i$th Dequeue in $b$th block of \nf{n} in \nf{n.parent} and then computes the order in $D(n.parent, superblock)$. Then calls \nf{index()} on \nf{n.parent} recursively. It is easy to see why the second is correct. Correctness of computing superblock comes from Lemma \ref{superBlock}.
%\end{proof}

\begin{lemma}[Computing SuperBlock]\label{superBlock}
  If \nf{Index(n,b,i)} performs line 101, then \nf{superblock} contains \nf{i}th Dequeue in \nf{b}th block of node \nf{n}.
\end{lemma}
%\begin{proof}
%\begin{enumerate}
% \item Value read for \nf{super[b.time]} in line 101 is not null.\\$\blacktriangleright$
%  Values \nf{c\textsubscript{dir}} read in lines 23, \nf{super} are set before incrementing in lines 26,27.
%
% \item \nf{super[]} preserves order from child to parent; if in a child block \nf{b} is before \nf{c} then \nf{b.time} $\leq$ \nf{c.time} and \nf{super[b.time]} $\leq$ \nf{super[c.time]}\\$\blacktriangleright$
%  Follows from the order of lines 60, 48, 49.
%
%% \item In a propagate step at most 2 different time values are read \\ If there are more than 2 numbers then the smallest number should have been propagated far before.
%% \item There are at most $p^2$ blocks with same time value in a node. \\ At most p processes could die before line 27 and each contains at most p elements.
% \item \nf{super[i+1]-super[i]}$\leq p$
%\\$\blacktriangleright$
% In a Refresh with successful CAS in line 46, \nf{super} and \nf{counter} are set for each child in lines 48,49. Assume the current value of the counter in node \nf{n} is \nf{i+1} and still \nf{super[i+1]} is not set. If an instance of successful \nf{Refresh(n)} finishes \nf{super[i+1]} is set a new value and a block is added after \nf{n.parent[sup[i]]}. There could be at most $p$ successful unfinished concurrent instances of \nf{Refresh()} that have not reached line 49. So the distance between \nf{super[i+1]} and \nf{super[i]} is less than $p$.
%
% \item Superblock of \nf{b} is within range $\pm 2p$ of the \nf{super[b.time]}.
%\\$\blacktriangleright$
%\nf{super[i]} is the index of the superblock of a block containing block b, followed by Lemma \ref{superCounter}. It is trivial to see that \nf{n.super} and \nf{n.b.counter} are increasing. \nf{super(b)} is the real superblock of b. \nf{super(t]} is the index of the superblock of the last block with time \nf{t}. If \nf{b.time} is \nf{t} we have:
%$$super[t]-p\leq super[t-1]\leq super(t-1] \leq super(b) \leq super(t+1)\leq super(t+1]\leq super[t]+p$$
%
%\end{enumerate}
%\end{proof}

>>>>

\usepackage{amsmath,amssymb,amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

@@

\Function{void}{Propagate()}{}
\If{\bf{not} \Call{Refresh()}{}} \label{firstRefresh}
\State \Call{Refresh()}{} \cmt{Lemma Double Refresh}\label{secondRefresh}
\EndIf
\If{\tt{this} \bf{is not} \tt{root}}
\State \tt{parent.}\Call{Propagate()}{}
\EndIf
\EndFunction{Propagate}

@@

\Function{boolean}{Refresh()}{}
\State \tt{s= size} \label{readSize}
%\If{\tt{n.blocks[h]!=null?}} \tt{h+=1}\EndIf
\State \tt{<new, np\sub{left}, np\sub{right}>= \Call{CreateBlock}{h}} \cmt{\tt{np\sub{left}, np\sub{right}} are the values read from the children's \tt{num\sub{propagated}} feild.}
\If{\tt{new.num==0}} \Return{\tt{true}} \cmt{The block contains nothing.}
\ElsIf{\tt{blocks.tryAppend(new, s)}}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{okcas}
\State \tt{\Call{CAS}{dir.super[np\sub{dir}], null, h+1}} \cmt{Write would work too.}
\State \tt{\Call{CAS}{dir.num\sub{propagated}, np\sub{dir}, np\sub{dir}+1}}
\EndFor
\State \tt{\Call{CAS}{size, h, h+1}}
\State \Return{\tt{true}}
\Else
\State \tt{\Call{CAS}{size, h, h+1}} \cmt{Even if another process wins, help to increase the size. The winner might have fallen sleep before increasing \tt{size}.}
\State \Return{ \tt{false}}
\EndIf
\EndFunction{Refresh}

@@

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{blocks[b].num\sub{enq}$\geq$i}.}
\Function{<int, int>}{IndexDeq}{\sl{int} b, \sl{int} i} \cmt{Returns the rank of $i$th dequeue in the $b$th block of the node, among the dequeues in the root.}
\If{\tt{this} \bf{is} \tt{root}}
\State\Return \tt{<b, i>}
\Else
\State \tt{dir= (parent.left==n)? left: right} \cmt{check if a left or a right child}
\State \tt{superBlock= parent.\Call{BSearch}{sum\sub{deq-dir}, i, super[blocks[b].group]-p, super[blocks[b].group]+p}} \label{computeSuper}
\Statex\cmt{superblock's group has at most $p$ difference with the value stored in \tt{super[]}.}
\If{\tt{dir {\keywordfont is} right}}
%\State \tt{i+= parent.blocks[superBlock-1].sum\sub{deq-right}}
\State \tt{i+= blocks[superBlock].sum\sub{deq-left}} \cmt{consider the dequeues from the right child}
\EndIf
\State \Return\tt{this.parent.}\Call{IndexDeq}{superBlock, i}
\EndIf
\EndFunction{Index}

@@

\framebox[1.1\width]{TEST} As a temporary test I have changed the name of n.size to n.established here, other options are n.head and n.lastBLock but they might be confusing since we have used them before. Fix the logical order of definitions (cyclic refrences).

\begin{definition}[Block]
A block is an object that stores some statistics described in Algorithm Queue. It implicitly shows a set of operations. The set of operations of block $b$ are the operations in the leaf subblocks of $b$. We show the set of operations of block $b$, set of blocks $B$ by $ops(b)$, $ops(B)$. We also say $b$ contains $op$ if $op\in ops(b)$.
\end{definition}

\begin{definition}[Order]
If \nf{n.blocks[i]==b}  we call \nf{i} the \emph{index} of block \nf{b}. Block \nf{b} is before block \nf{b$^\prime$} in node \nf{n} if and only if \nf{b}'s index is smaller than \nf{b$^\prime$}'s. 
\end{definition}

\begin{definition}[Subblock]
Block \texttt{b} is a subblock of \texttt{n.blocks[i]} if it is $\in$ \texttt{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} $\cup$ \texttt{n.right.blocks[n.blocks[i-1].end\textsubscript{right}+1..n.blocks[i].end\textsubscript{right}]} or  is a subblock of a block in it.
\end{definition}

For simplicity we say block \nf{b} is propagated to node \nf{n} or set \nf{S} if \nf{b} is in \nf{n.blocks} or \nf{S} or is a subblock of a block in \nf{n.blocks} or \nf{S}.

@@

\begin{definition}
  Block \nf{b} in \nf{n.blocks} is \emph{Established} at time $t$ if \nf{n.established} is greater than index of \nf{b} at time $t$. Block \nf{b} in \nf{n.blocks} is in \emph{EST\sub{n, t}} if \nf{b} is a subblock of \nf{b$^\prime$} in \nf{n.blocks} such that \nf{b$^\prime$} is established  at time $t$.
\end{definition}

\begin{lemma}[headProgress] \label{lem::sizeProgress}
\nf{n.established} is non-decreasing over time.
\end{lemma}
%\begin{proof}
%  Simply because \nf{n.head} is only incremented.
%\end{proof}

\begin{lemma}[headPosition] \label{lem::headPosition} The value read in Line \ref{prevLine}(\nf{h=n.established}) might be 1 bit behind the first empty block in the node.
\end{lemma}
\begin{proof}
 Because at the end of every \nf{Refresh()} with block size greater than 0 (Lines 53,56) \nf{n.head} is incremented. Maybe some process goes to sleep before incrementing the head, but after sleeping if \nf{h} does not increase then \nf{CAS} in Line 52 is going to be failed and nothing is going to be appended to \nf{n.blocks}.
\end{proof}

@@

\begin{lemma}[establishedOrder]\label{lem::establishedOrder}
  If  time $t<$ time $t^\prime$, then \nf{ops(EST\sub{n, t})}$\subseteq$ \nf{ops(EST\sub{n, t$^\prime$})}.
\end{lemma}
\begin{proof}
Because blocks are only appended(not modified) with CAS to \nf{n.blocks[n.head]} and \nf{n.head} is non-decreasing.
\end{proof}

\begin{lemma}[createBlock] \label{lem::createBlock}
  If \nf{b} is the block returned by \nf{n.CreateBlock(h, x)} invoked at time $t$, then \nf{ops(EST\sub{n.left, t})} $\cup$ \nf{ops(EST\sub{n.right, t})} - \nf{ops(EST\sub{n, t})} $\subseteq$ \nf{ops(b)}.
\end{lemma}
\begin{proof}
 We prove the claim for the left child. Blocks in \nf{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} are all the new established operations at time $t$ by definition of Subblock. Line 70 is after $t$ and since the head is only increasing (Lemma~\ref{lem::headProgress}) the lemma holds. See Figure~\ref{fig::createBlock}. The right child is the same.
\end{proof}

\begin{lemma}[trueRefresh] \label{lem::trueRefresh}
  Let $t_i$ be the time \nf{n.Refresh()} is invoked and $t_t$ be the time it is terminated. Suppose \nf{n.Refresh()}'s \nf{TryAppend(new, s)} returns \nf{true}, then \nf{ops(EST\sub{n.left, t\sub{i}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}})} $\in$ \nf{ops(EST\sub{n, t\sub{t}})}.
\end{lemma}
\begin{proof}
By Lemma~\ref{lem::createBlock} \nf{new} contains \nf{n}'s childrens' established blocks before Line 43 which is  appended to \nf{n.blocks} by \nf{CAS} in Line 48.
\end{proof}

\begin{lemma}[falseRefresh] \label{lem:falseRefresh}
    If instance \nf{r} of \nf{n.Refresh()} reads value \nf{s} on line \ref{readSize} and then returns \nf{false}, then there is another instance $r^\prime$ of \nf{n.Refresh()} that has performed a successful \nf{TryAppend(new, s)}. A \nf{TryAppend()} is successful if its \nf{CAS} is successful.
\end{lemma}
\begin{proof}
  If there is no other concurrent successful \nf{Refresh(n)} then \nf{Refresh(n)} would succeed in Line 48. So there is another \nf{Refresh(n)}, that has to \nf{CAS}ed successfully its block in \nf{n.blocks[h]} after Line 43 of \nf{Refresh(n)}. Otherwise the other \nf{Refresh(n)} should have read \nf{h$^\prime$>h} instead of \nf{h} for \nf{n.head}(Line 52).
\end{proof}

\begin{lemma}[Double Refresh] \label{doubleRefresh}
  Consider two consecutive failed instances $r_1$,$r_2$ of \nf{n.Refresh()} by some process. Let $t_1$ be the time $r_1$ is invoked and $t_2$ be the time $r_2$ terminated. After $r_2$'s \nf{TryAppend} we have \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{2}})}.
\end{lemma}
\begin{proof}

If Line 35 (first Refresh which we call $R_1$) returns \nf{true}, the claim is held by Lemma~\ref{lem::trueRefresh}. If not, then there is another successful instance of \nf{Refresh()} $R_1^\prime$ by Lemma~\ref{lem:falseRefresh}. $R_1^\prime$ may or may not have propagated some subblocks of \nf{new} to \nf{n}. It is obvious that the \nf{new} constructed by the second Refresh in Line 36 contains the blocks in \nf{new} by $R_1$ which $R_1^\prime$ did not contain, since \nf{n.head} is only increasing (Maybe Lemma~\ref{lem::oldnewOrder}). If $R_2$ succeeds by Lemma~\ref{lem::trueRefresh} the claim holds. If not, it is deduced that \nf{n.blocks[h]} was not \nf{null} before $R_2$'s \nf{CAS}. Furthermore, \nf{n.blocks[h]} was \nf{null} before reading \nf{h} by $R_1$. So there is a successful \nf{Refresh()} after the read of \nf{h} in $R_1$ and before the \nf{CAS} of $R_2$. This \nf{Refresh()} contains all the new established operations before Line 35 (Maybe Lemma~\ref{lem::oldnewOrder}) and by Lemma~\ref{lem::trueRefresh} our claim holds. See Figure~\ref{}.
\end{proof}

\begin{figure}[hbt]
  \center\includegraphics[width=3in]{pics/doublyrefresh.drawio.png}
  \caption{$R_2^\prime$'s \nf{CAS} is executed after \nf{h1=n.head}.}
\end{figure}

%  Block \nf{new} is created of new established subblocks of children of \nf{n}(Lemma \ref{lem::createBlock}, Line 46). If \nf{CAS} in Line 48 succeeds then by Lemma~\ref{lem::trueRefresh} new established blocks will be in \nf{n}.

\begin{corollary}[Propagate Step] \label{doublyRefresh}
All operations in \nf{n}'s children's established blocks before line \ref{firstRefresh} are guaranteed to be in \nf{n}'s blocks after line~\ref{secondRefresh}.
\end{corollary}
\begin{proof}
Suppose block \nf{b} with index \nf{i} is in the the left child of \nf{n} before the line 35. By Lemma~\ref{head} it follows that \nf{n.left.head} is greater than \nf{i}. \nf{Refresh()} calls \nf{CreateBlock()} and creates a block from blocks between \nf{n.blocks[n.head].end\textsubscript{left}} and \nf{n.left.head} in the left child, which contains \nf{b} as well. First it tries to append it in \nf{n.blocks.head} and if it was succsuful it continues recursively. If not it tries again, and if the second call of \nf{Refresh()} in Line 36 fails. It means there is another \nf{Refresh} which has reah its \nf{i} after the Line 35, so it contains \nf{b} as well.
\end{proof}
\nf{CreateBlock()} reads blocks in the children that do not exist in the parent and aggregates them into one block. If a \nf{Refresh()} procedure returns true it means it has appended the block created by \nf{CreateBlock()} into the parent node's sequence. So suppose two \nf{Refresh}es fail. Since the first \nf{Refresh()} was not successful, it means another CAS operation by a \nf{Refresh}, concurrent to the first \nf{Refresh()}, was successful before the second \nf{Refresh()}. So it means the second failed \nf{Refresh} is concurrent with a successful \nf{Refresh()} that assuredly has read block before the mentioned line \nf{35}. After all it means if any of the \nf{Refresh()} attempts were successful the claim is true, and also if both fail the mentioned claim still holds.

\begin{lemma}[No Duplicates]\label{append}
If \nf{op} is appended to \nf{n.blocks[i]} then after that there is no \nf{j>i} such that \nf{op}$\in$\nf{ops(n.blocks[j])}.
\end{lemma}
\begin{proof}
\nf{Append(op)} adds \nf{op} to \nf{l\textsubscript{p}.blocks}(Line~\ref{addOP}) and \nf{Propagate()} recursively propagates \nf{op} up to the root.
By lemma \ref{doublyRefresh} we know that operation \nf{op} propagates from child to parent at each level.
\end{proof}
\begin{corollary}
  After \nf{Append(blk)} finishes \nf{ops(blk)}$\subseteq$\nf{ops(root.blocks[x])} for some \nf{x} and only one \nf{x}.
\end{corollary}

\begin{lemma}[Block Size Upper Bound]\label{blockSize}
Each block contains at most one operation from each processs \nf{($\forall$ \nf{process p:}\#\nf{operations of p}$\in$\nf{ops(n.blocks[x]$\leq$1})}.
\end{lemma}
%\begin{proof}
%  Note that \nf{prevIndex} and \nf{lastIndex} defined in lines \ref{lastLine}, \ref{prevLine} are the indices defined in the Definition~\ref{def::subblock}. After a block has propagated to the parent the blocks between \nf{prevIndex} and \nf{lastIndex} make up to the parent (Lemma~\ref{doublyRefresh}). The number of new operations which have not propagated yet to the parent cannot be more than $p$. If so by the law of pigeonholes there is a process which has appended two concurrent operations.
%\end{proof}

\begin{lemma}[Subblocks Upperbound]\label{subBlocksBound}
Each block has at most $p$ subblocks.
\end{lemma}
%\begin{proof}
%  From Line~\ref{addOP} it is induced directly that each block contains at least one operation. Now it follows directly by Lemma \ref{blockSize}.
%\end{proof}

\pagebreak
\begin{definition} [Ordering of operations inside the nodes] \label{ordering}
$\blacktriangleright$ Note that from Lemma \ref{blockSize} we know there is at most one operation from each process in a given block.

\begin{itemize}
\item We call operations before $op$ in the sequence of operations $S$, prefix of the $op$.
  \item $E(n,i)$ is the sequence of enqueue operations $\in$ \nf{ops(n.blocks[i])} ordered by their process id.
  \item $D(n,i)$ is the sequence of dequeue operations $\in$ \nf{ops(n.blocks[i])} ordered by their process id.
\item Order of the enqueue operations in $n$: $E(n)=E(n,1).E(n,2).E(n,3)...$
\item Order of the dequeue operations in $n$: $D(n)=D(n,1).D(n,2).D(n,3)...$
\item Linearization: $L=E(root,1).D(root,1).E(root,2).D(root,2).E(root,3).D(root,3)...$
\end{itemize}
\end{definition}
\it{Note that in the non-root nodes we only order enqueues and dequeues among the operations of their own type. Since \nf{GetENQ()} only works on enqueues and \nf{IndexDEQ()} works on deqeueus.}

@@

\begin{lemma}[Get corretness] \label{get}
\nf{n.GetENQ(b,i)} returns the $i$th Enqueue in $E(n,b)$.
\end{lemma}
%\begin{proof}
%It is obvious that \nf{Get(leaf l,b,1)} returns the operations stored in $b$th block of leaf $l$. To find the \nf{i}th enqueue in block \nf{b} of an internal node \nf{n} in line 87 it is decided that it resides in the left child or the right child. This decision is made by Definition of $E(n,b)$. After that Lines 88, 92 search the proper subblocks of \nf{b}. From Definition~\ref{def::subblock} weknow the subblocks of the $b$th block are within the \nf{prevBlock} and \nf{lastBlock} block of the \nf{CreateBlock()}.
%\end{proof}

\begin{lemma}[Index correctness]
 \nf{n.Index(b,i)} returns the rank in the $D(root)$ of ith Dequeue in $D(n,b)$.
\end{lemma}
%\begin{proof}
%  \nf{Index(n,b,i)} computes superblock of $i$th Dequeue in $b$th block of \nf{n} in \nf{n.parent} and then computes the order in $D(n.parent, superblock)$. Then calls \nf{index()} on \nf{n.parent} recursively. It is easy to see why the second is correct. Correctness of computing superblock comes from Lemma \ref{superBlock}.
%\end{proof}

\begin{lemma}[Computing SuperBlock]\label{superBlock}
After computing line \ref{computeSuper} of \nf{n.IndexDEQ(b,i)}, \nf{superblock} contains the \nf{i}th dequeue in the \nf{b}th block of the node \nf{n}.
\end{lemma}
%\begin{proof}
%\begin{enumerate}
% \item Value read for \nf{super[b.time]} in line 101 is not null.\\$\blacktriangleright$
%  Values \nf{c\textsubscript{dir}} read in lines 23, \nf{super} are set before incrementing in lines 26,27.
%
% \item \nf{super[]} preserves order from child to parent; if in a child block \nf{b} is before \nf{c} then \nf{b.time} $\leq$ \nf{c.time} and \nf{super[b.time]} $\leq$ \nf{super[c.time]}\\$\blacktriangleright$
%  Follows from the order of lines 60, 48, 49.
%
%% \item In a propagate step at most 2 different time values are read \\ If there are more than 2 numbers then the smallest number should have been propagated far before.
%% \item There are at most $p^2$ blocks with same time value in a node. \\ At most p processes could die before line 27 and each contains at most p elements.
% \item \nf{super[i+1]-super[i]}$\leq p$
%\\$\blacktriangleright$
% In a Refresh with successful CAS in line 46, \nf{super} and \nf{counter} are set for each child in lines 48,49. Assume the current value of the counter in node \nf{n} is \nf{i+1} and still \nf{super[i+1]} is not set. If an instance of successful \nf{Refresh(n)} finishes \nf{super[i+1]} is set a new value and a block is added after \nf{n.parent[sup[i]]}. There could be at most $p$ successful unfinished concurrent instances of \nf{Refresh()} that have not reached line 49. So the distance between \nf{super[i+1]} and \nf{super[i]} is less than $p$.
%
% \item Superblock of \nf{b} is within range $\pm 2p$ of the \nf{super[b.time]}.
%\\$\blacktriangleright$
%\nf{super[i]} is the index of the superblock of a block containing block b, followed by Lemma \ref{superCounter}. It is trivial to see that \nf{n.super} and \nf{n.b.counter} are increasing. \nf{super(b)} is the real superblock of b. \nf{super(t]} is the index of the superblock of the last block with time \nf{t}. If \nf{b.time} is \nf{t} we have:
%$$super[t]-p\leq super[t-1]\leq super(t-1] \leq super(b) \leq super(t+1)\leq super(t+1]\leq super[t]+p$$
%
%\end{enumerate}
%\end{proof}
